Zero-shot task adaptation by homoiconic meta-mapping

### 2019-09-23 writing/homoiconic_meta_mapping_main.tex
Tweaks
Reframing intro/conclusions slightly

\begin{abstract}
How can deep learning systems flexibly reuse their knowledge? Toward this goal, we propose a new class of challenges, and a class of architectures that can solve them. The challenges are meta-mappings, which involve systematically transforming task behaviors to adapt to new tasks zero-shot. We suggest that the key to achieving these challenges is representing the task being performed along with the computations used to perform it. We therefore draw inspiration from functional programming and recent work in meta-learning to propose a class of Homoiconic Meta-Mapping (HoMM) approaches that represent data points and tasks in a shared latent space, and learn to infer transformations of that space. HoMM approaches can be applied to any type of machine learning task, including supervised learning and reinforcement learning. We demonstrate the utility of this perspective by exhibiting zero-shot remapping of behavior to adapt to new tasks.
\end{abstract}

%\vspace{-0.75em} % section
\section{Introduction}
%\vspace{-0.5em} % section
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
\looseness=-1
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[see also section \ref{sec_discussion}]{Santoro2016, Finn2017a, Finn2018, Stadie2018, Botvinick2019}. However, humans can use our knowledge of a task to flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose using architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing task demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. By allowing the network to recursively treat its task representations as data points, and transform them to produce new task representations, our approach is able to achieve this flexibility parsimoniously. We suggest that approaches like ours will be key to building more intelligent and flexible deep learning systems. \par

%\vspace{-0.25em}
\section{Meta-mapping}
%\vspace{-0.5em} % section
%Can cut "the task"s in go sentence if needed
We propose the meta-mapping challenge. We define a meta-mapping as a task, or mapping, that takes a task as an input, output, or both. These include mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. (We discuss the relationship between our work and prior work in section \ref{sec_discussion}.) This adaptation can be cued in several ways, including examples of the mapping (after winning and losing at poker, try to lose at blackjack) or natural-language instructions (``try to lose at blackjack''). \par
We argue that task-to-task meta-mappings are a useful way to think about human-like flexibility, because a great deal of our rapid adaptation is from a task to some variation on that task. For example, the task of playing go on a large board is closely related to the task of playing go on a small board. Humans can exploit this to immediately play well on a different board, but deep learning models generally have no way to achieve this. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose, or trying to achieve some orthogonal goal. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not fully capture human flexibility. Given appropriate conditioning (see below), our architecture can adapt to substantial task alterations zero-shot, that is, without seeing a single example from the new task \citep{Lake2016}. We suggest that meta-mappings offer a way to understand this flexibility, and a way for deep-learning models to achieve it. Achieving the flexibility to adapt to meta-map to new tasks will be an important step toward more general intelligence -- intelligence that is not limited to precisely the training domains it has seen. \par

@@

For example, suppose we have an embedding $z_{game1} \in Z$ for the task of playing some game. We can generate a meta-mapping embedding $z_{\text{meta}} \in Z$ from embeddings generated by the system when it is trying to win and lose various games: $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{game2},z_{game2,lose}\right), ... \right\}\right)$. We can generate a new task embedding $\hat{z}_{new} \in Z$:  
\[\hat{z}_{new} = F_{z_{\text{meta}}}(z_{game1}) \qquad \text{where } F_{z_{\text{meta}}} \text{ is parameterized by } \mathcal{H}\left(z_{\text{meta}}\right)\]

@@

This $\hat{z}_{new}$ could be interpreted as the system's guess at a losing strategy for game 1. To train a meta-mapping, we minimize the $\ell_2$ loss in the latent space betwen this guessed embedding and the embedding of the target task. Whether or not we have such a target embedding, we can evaluate how well the system loses with this $\hat{z}_{new}$ strategy, by stepping back down a level of abstraction and actually having it play the game via this embedding (fig. \ref{architecture_inference_fig}c). This is how we evaluate meta-mapping performance -- evaluating the loss of transformed task embeddings on the respective target tasks. \par
Alternatively, we could map from language to a meta-mapping embedding, rather than inducing it from examples of the meta-mapping. This corresponds to the human ability to change behavior in response to instructions. The key feature of our architecture -- the fact that tasks, data, and language are all embedded in a shared space -- allows for substantial flexibility within a unified framework. Furthermore, our approach is parsimonious. Because it uses the same meta-learner for both basic tasks and meta-mappings, this increased flexibility does not require any added parameters.\footnote{At least in principle, in practice of course increasing network size might be more beneficial for HoMM architectures performing meta-mappings as well as basic tasks than for those performing only basic tasks.}

@@

\caption{From language input.}

@@

\caption{From language input.}

@@

We've highlighted a new type of flexibility in the form of meta-mapping between tasks. Meta-mapping can produce zero-shot performance on novel tasks, based on their relationship to old tasks. However, achieving this flexibility is difficult in architectures that separate computation from the knowledge of what is being computed. To address this, our Homoiconic Meta-Mapping (HoMM) approach explicitly represents tasks by function embeddings, and derives the computation from these embeddings via a HyperNetwork. This allows for our key insight: by embedding tasks and individual data points in a shared latent space, the same meta-learning architecture can be used both to transform data for basic tasks, and to transformat task embeddings to adapt to task variations. This approach is parsiomious, because it uses the same networks for basic- and meta-mappings. Perhaps because of this, the requisite meta-task sample complexity is small; we showed generalization to unseen meta-mappings after training on only 20 meta-mappings. That is, we are able to achieve zero-shot performance on a novel task based on a held-out relationship between tasks. This is a step closer to human-level flexibility.\par

>>>>

\begin{abstract}
How can deep learning systems flexibly reuse their knowledge? Toward this goal, we propose a new class of challenges, and a class of architectures that can solve them. The challenges are meta-mappings, which involve systematically transforming task behaviors to adapt to new tasks zero-shot. We suggest that the key to achieving these challenges is representing the task being perform in such a way that this representation is itself transformable. We therefore draw inspiration from functional programming and recent work in meta-learning to propose a class of Homoiconic Meta-Mapping (HoMM) approaches that represent data points and tasks in a shared latent space, and learn to infer transformations of that space. HoMM approaches can be applied to any type of machine learning task, including supervised learning and reinforcement learning. We demonstrate the utility of this perspective by exhibiting zero-shot remapping of behavior to adapt to new tasks.
\end{abstract}

%\vspace{-0.75em} % section
\section{Introduction}
%\vspace{-0.5em} % section
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep{Lake2016, Marcus2018}. The problem of rapid learning has been partially addressed by meta-learning systems \citep[see also section \ref{sec_discussion}]{Santoro2016, Finn2017a, Finn2018, Stadie2018, Botvinick2019}. However, humans can use our knowledge of a task to flexibly adapt when the task changes. In particular, we can often perform an altered task zero-shot, that is, without seeing any data at all. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can perform reasonably on our first attempt. \par
One fundamental reason for this is that humans are aware of what we are trying to compute and why. This allows us to adapt our task representations to perform a new task zero-shot. By contrast, most deep learning models do not explicitly represent tasks at all. Those that do (such as some meta-learning models) generally do not represent relationships between tasks, and so cannot adapt to a new task without data from the new task. To address this, it is necessary to represent tasks in a transformable space. This can grant the ability to rapidly adapt behavior to a new task. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). As noted above, this type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to deep-learning models. To address this challenge, we propose using architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing task demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. By allowing the network to recursively treat its task representations as data points, and transform them to produce new task representations, our approach is able to achieve this flexibility parsimoniously. We suggest that approaches like ours will be key to building more intelligent and flexible deep learning systems. \par

%\vspace{-0.25em}
\section{Meta-mapping}
%\vspace{-0.5em} % section
%Can cut "the task"s in go sentence if needed
We propose the meta-mapping challenge. We define a meta-mapping as a task, or mapping, that takes a task as an input, output, or both. These include mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. (We discuss the relationship between our work and prior work in section \ref{sec_discussion}.) This adaptation can be cued in several ways, including examples of the mapping (after winning and losing at poker, try to lose at blackjack) or natural-language instructions (``try to lose at blackjack''). \par
We argue that task-to-task meta-mappings are a useful way to think about human-like flexibility, because a great deal of our rapid adaptation is from a task to some variation on that task. For example, the task of playing go on a large board is closely related to the task of playing go on a small board. Humans can exploit this to immediately play well on a different board, but deep learning models generally have no way to achieve this. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose, or trying to achieve some orthogonal goal. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not fully capture human flexibility. Given appropriate conditioning (see below), our architecture can use meta-mappings to adapt to substantial task alterations zero-shot, that is, without seeing a single example from the new task \citep{Lake2016}. Achieving this flexibility to meta-map to new tasks will be an important step toward more general intelligence -- intelligence that is not limited to precisely the training tasks it has seen. \par

@@

For example, suppose we have an embedding $z_{game1} \in Z$ for the task of playing some game, and we want to switch to trying to lose this game. We can generate a meta-mapping embedding $z_{\text{meta}} \in Z$ from examples of embeddings generated by the system when it is trying to win and lose various games: $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{game2},z_{game2,lose}\right), ... \right\}\right)$. We can generate a new task embedding $\hat{z}_{game1,lose} \in Z$:  
\[\hat{z}_{game1,lose} = F_{z_{\text{meta}}}(z_{game1}) \qquad \text{where } F_{z_{\text{meta}}} \text{ is parameterized by } \mathcal{H}\left(z_{\text{meta}}\right)\]

@@

This $\hat{z}_{games1,lose}$ can be interpreted as the system's guess at a losing strategy for game 1. To train a meta-mapping, we minimize the $\ell_2$ loss in the latent space betwen this guessed embedding and the embedding of the target task\footnote{The gradients do not update the example function embeddings, only the weights of \(\mathcal{M}\) and \(\mathcal{H}\), due to memory contraints. Allowing this might be useful in more complex applications.}. Whether or not we have such a target embedding, we can evaluate how well the system loses with this $\hat{z}_{game1,lose}$ strategy, by stepping back down a level of abstraction and actually having it play the game via this embedding (fig. \ref{architecture_inference_fig}c). This is how we evaluate meta-mapping performance -- evaluating the loss of transformed task embeddings on the respective target tasks. \par
Alternatively, we could map from language to a meta-mapping embedding, rather than inducing it from examples of the meta-mapping. This corresponds to the human ability to change behavior in response to instructions. The key feature of our architecture -- the fact that tasks, data, and language are all embedded in a shared space -- allows for substantial flexibility within a unified framework. Furthermore, our approach is parsimonious. Because it uses the same meta-learner for both basic tasks and meta-mappings, this increased flexibility does not require any added parameters.\footnote{At least in principle, in practice of course increasing network size might be more beneficial for HoMM architectures performing meta-mappings as well as basic tasks, compared to those performing only basic tasks.}

@@

\caption{The polynomials domain.}

@@

\caption{The cards domain.}

@@

We've highlighted a new type of flexibility in the form of meta-mapping between tasks. Meta-mapping can produce zero-shot performance on novel tasks, based on their relationship to old tasks. This is a key aspect of human flexibility, but is lacking from most contemporary deep-learning models. Achieving this flexibility requires representing tasks in a transformable way. To address this, our Homoiconic Meta-Mapping (HoMM) approach explicitly represents tasks by function embeddings, and derives the computation from these embeddings via a HyperNetwork. This allows for our key insight: by embedding tasks and individual data points in a shared latent space, the same meta-learning architecture can be used both to transform data for basic tasks, and to transform task embeddings to adapt to task variations. This approach is parsiomious, because it uses the same networks for basic- and meta-mappings. Perhaps because of this, the requisite meta-task sample complexity is small; we showed generalization to unseen meta-mappings after training on only 20 meta-mappings. That is, we are able to achieve zero-shot performance on a novel task based on a held-out relationship between tasks. This is a step closer to human-level flexibility.\par
